---
title: "2_Avocados"
author: "Jeremy Whorton"
date: "1/29/2019"
output: html_document
---

## Avocado Sales

In this lab we’re going to be looking at avocado prices! The dataset comes to us from kaggle and represents weekly retail scan data: avocado.csv. A description of the data can be found at the Hass Avocado Board website.

```{r read data/setup}
library(tidyverse)
avo <- read.csv("https://www.dropbox.com/s/vsc1dkosz6nwake/avocado.csv?dl=1")
```

1) Which region sold the most bags of small organic avocados in 2017?
```{r region small bags}
avo %>%
  group_by(region) %>%
  filter(region != "TotalUS") %>%
  filter(type == "organic") %>%
  summarize(Small.Bags = sum(Small.Bags)) %>%
  top_n(1)
```

2) Use separate() to split the Date variable into year, month, and day. In which month is the highest volume of avocado sales?
```{r month volume}
avo %>%
  separate(Date, sep = "-", into = c("Year", "Month", "Day")) %>%
  group_by(Month) %>%
  summarize(Total.Volume = sum(Total.Volume)) %>%
  top_n(1)
```


3) Which regions sell the most avocados by volume? Plot side-by-side boxplots of Total Volume for only the 5 regions with the highest averages for the Total Volume variable.
```{r region sales plot}
top_regions <- avo %>%
  group_by(region) %>%
  filter(region != "TotalUS") %>%
  summarize(Total.Volume = sum(Total.Volume)) %>%
  top_n(5)

avo %>%
  filter(region == top_regions$region) %>%
  ggplot(aes(x = region, y = Total.Volume)) + geom_boxplot()
```



4) In which regions is the price of organic versus conventional avocados most different? Support your answer with a few summary numbers and a plot.
```{r organic}
regional_types <- avo %>%
  group_by(region, type) %>%
  summarize(AveragePrice = mean(AveragePrice))
  
regional_types %>%
  mutate(difference = AveragePrice - lag(AveragePrice)) %>%
  na.omit() %>%
  summarize(difference = max(difference)) %>%
  top_n(5)

regional_types %>%
  ggplot(aes(x = type, y = AveragePrice, color = region, group = region)) + geom_point() + geom_line()
```
In the graph, the lines with the steepest slope have the highest differences between conventional and organic prices. 


5) How do their avocado habits differ? Make a plot that shows, for all 4 California regions, the percent of avocado sales that are small, large, or extra large. Separate your plot by conventional vs. organic avocados.
```{r size in california}
habits <- avo %>%
  filter(region == "LosAngeles" | region == "Sacramento" | region == "SanDiego" | region == "SanFrancisco") %>%
  group_by(region, type) %>%
  summarize(large_percent = sum(Large.Bags) / sum(Total.Bags),
         medium_percent = sum(Small.Bags) / sum(Total.Bags),
         xlarge_percent = sum(XLarge.Bags) / sum(Total.Bags)) %>%
  gather(key = size, value = percent)
habits
 habits %>%
   ggplot(aes(x = region, y = percent, fill = size)) + geom_col() + facet_grid(. ~ type)

# habits %>%
#   summarize(xlarge_percent = mean(xlarge_percent)) %>%
#   ggplot(aes(x = region, y = xlarge_percent)) + geom_point() + facet_grid(. ~ type)
# habits %>%
#   summarize(medium_percent = mean(medium_percent)) %>%
#   ggplot(aes(x = region, y = medium_percent)) + geom_point() + facet_grid(. ~ type)
#   
```
 