---
title: "8_Activity_Testing"
author: "Jeremy Whorton"
date: "February 6, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(broom)

data("starwars")
starwars %>%
  head()
```

# Instructions

The dataset "starwars" has been loaded for you.  Take a look at it and understand the variables. 

## Writing your own t-test

Write a function called `run_t_test` that takes as input two vectors of numbers, and an option for the alternative hypothesis, and gives back the results of a t-test for equality of means.  You may not use the function `t.test()` except to check that your version is working correctly.

```{r}
run_t_test <- function(x, y, alternative = "two.sided"){
  
  x <- x[!is.na(x)]
  y <- y[!is.na(y)]
  
  x_bar <- mean(x)
  y_bar <- mean(y)
  
  nx <- length(x)
  ny <- length(y)
  
  sdx <- sd(x)
  sdy <- sd(y)
  
  se = sqrt(sdx^2/nx + sdy^2/ny)
  t = (x_bar - y_bar) / se
  df = min(nx, ny)
  
  p_val <- 2*pt(-abs(t), df = df)
  
  if (alternative == "less") {
    p_val <- pt(-t, df = df)
  } else if (alternative == "greater") {
    p_val <- 1 - pt(t, df = df)
  }
  
  return(list(test.stat = t, df = df, p.val = p_val))
}
bob <- c(1:15, NA)
joe <- c(1:20, NA)

run_t_test(bob, joe, "more")
```

Use your t-test function to answer the following:

1. Are Humans or Droids taller?
```{r}
starwars_h_d <- starwars %>%
  filter(species == "Droid" | species == "Human") %>%
  select(species, height) %>%
  group_by(species) %>%  
  mutate(row_id=1:n()) %>% 
  ungroup() %>%  
  spread(key=species, value=height) %>%    
  select(-row_id) 
run_t_test(starwars_h_d$Human, starwars_h_d$Droid, "more")
```
There is not signifcance evidence that humans are taller than droids, thus we fail to reject the null.

2. Do men or women weigh more?
```{r}
starwars_m_w <- starwars %>%
  select(gender, mass) %>%
  group_by(gender) %>%   
  mutate(row_id=1:n()) %>% 
  ungroup() %>%  
  spread(key=gender, value=mass) %>%    
  select(-row_id) 
run_t_test(starwars_m_w$male, starwars_m_w$female, "more")
```
There is statistial evidence that males weigh more than females.

## Chi-square tests

Use the function `chisq.test()` to answer the following:

1. Is there a relationship between hair color and eye color?
```{r}
starwars_eyes <- starwars %>% 
  count(hair_color, eye_color) %>%
  spread(key = hair_color, value = n) %>%
  select(-eye_color)

starwars_eyes[is.na(starwars_eyes)] <- 0

starwars_eyes %>%
  chisq.test() %>%
  tidy()
```
Yes, absolutely there is a relationship between eyecolor and hair color.

2. Do Droids tend to come from some homeworlds more than others?
```{r}
starwars_dr <- starwars %>% 
  filter(species == "Droid") %>%
  count(species, homeworld) %>%
  spread(key = homeworld, value = n) %>%
  select(-species)

starwars_dr[is.na(starwars_dr)] <- 0

starwars_dr %>%
  chisq.test() %>%
  tidy()
```
No, there is little relatonship between the homeworld a droid comes from. 

## Many tests (optional)

On which worlds is the height difference between men and women most pronounced?
 



